A fast unified model for parsing and sentence understanding

### 2016-03-18 writing/hard_stack_paper/paper.tex
Major cleanup/trimming/overhaul.

A wide range of current state-of-the-art models in NLP are built around neural network components that build vector representations of sentence meaning \citep{socher2011semi,sutskever2014sequence}. This component, the sentence encoder, is generally formulated as a learned parametric function from a sequence of word vector to a sentence vector, and this function can take a range of different forms. Common sentence encoders include sequence-based recurrent neural network models (RNNs) with Long Short-Term Memory \citep[LSTM,][see Figure~\ref{fig:batching:good}]{hochreiter1997long}, which accumulate information over the sentence sequentially, convolutional neural networks over words or characters \citep{kalchbrenner2014convolutional,DBLP:journals/corr/ZhangZL15}, which accumulate information using filters over short local sequences of words, and tree-structured recursive neural networks \citep[TreeRNNs,][see Figure~\ref{fig:batching:bad}]{goller1996learning,socher2011semi}, which propagate information up a binary parse tree structure over words.

Of these, the TreeRNN is the principled choice, since meaning in natural language sentences constructed incrementally according to a tree structure \citep[][i.a.]{Szabolcsi:2009}. However, while these models have shown promise \citep{tai2015improved,li2015tree,bowman2015trees}, they have largely been overlooked in favor of sequence-based RNNs because of their incompatibility with batched computation and their reliance on external parsers.  Batched computation---performing synchronized computation across many examples at once---yields order-of-magnitude improvements in model run time, and is a crucial enabling technology in allowing neural networks to be trained efficiently on large datasets. Because TreeRNNs use a different model structure for each sentence, batched computation is impossible in standard implementations. In addition, standard TreeRNN models can only operate on sentences that have already been processed by a syntactic parser, which slows and complicates the use of these models at test time for most applications.

This paper introduces a new model to address both these issues: the Stack-augmented Parser-Interpreter Neural Network, or SPINN, shown in Figure~\ref{fig:m1-views}. SPINN executes the computations of a tree-structured model in a linearized sequence, and can incorporate a neural network parser that determines the overall tree structure on-the-fly. This design improves upon the TreeRNN architecture in three ways: At test time, it can simultaneously parse and interpret unparsed sentences without incurring a substantial additional computational cost, removing the dependence on an external parser. In addition, it supports batched computation for both parsed and unparsed sentences, supporting dramatic speedups over standard TreeRNNs. Finally, it supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over either pure sequence-based models or pure tree-based models.

@@

\citet{socher2011semi} and \citet{socher2011parsing} present versions of the TreeRNN model which are capable of operating over unparsed inputs. However, these methods require an expensive search process at test time. Our model presents a fast alternative approach.

@@

SPINN is inspired by the shift-reduce parsing formalism \citep{aho1972theory}, which builds a tree structure over a sequence (e.g., a natural language sentence) by a single left-to-right scan over the tokens. Shift-reduce parsing is frequently used in natural language parsing \citep[e.g.,][]{nivre2003efficient}.

A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and uses transitions $\mathbf t = (t_0, \dots, t_{T-1})$, where each $t_i \in \{\shift, \reduce\}$ specifies a \shift~or \reduce~transition (described below). In general a parser may also generate these predictions on-the-fly as it reads the tokens. It reads left-to-right over the input transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure. For any binary-branching tree structure over $N$ words, this requires $T = 2N - 1$ transitions.

The parser uses two auxiliary data structures, a stack $S$ of partially completed subtrees and a buffer $B$ of tokens yet to be parsed. We initialize the parser with an empty stack and with a buffer containing all the sentence tokens $\mathbf x$. Let $\langle S, B \rangle = \langle \emptyset, \mathbf x \rangle$ denote this starting state. We next proceed through the transition sequence, where each transition $t_i$ selects one of the two following operations. Below the $\mid$ symbol denotes the \textit{cons} (concatenation) operator. We arbitrarily choose to always \textit{cons} on the left in the notation below.
\begin{description}
  \item[\shift:] $\langle S, x \mid B \rangle \to \langle x \mid S, B \rangle$. This operation pops an element from the buffer and pushes it onto the top of the stack.
  \item[\reduce:] $\langle x \mid y \mid S, B \rangle \to \langle (x, y) \mid S, B \rangle$. This operation pops the top two elements from the stack, merges them into a binary tree with children $(x, y)$, and pushes the result back onto the stack.
\end{description}

@@

The core SPINN model implements a sequential shift-reduce parser which operates over sentences. It is designed to produce a vector representation of the sentence as output, as opposed to a tree as in standard shift-reduce parsing. This relies on two key modifications to the shift-reduce algorithm: First, the parser's intermediate representations, including those of partial tree structures on the stack, are fixed-length vectors. In addition, the \reduce~operation combines two such representations of trees on the stack into a larger tree using a version of the TreeLSTM neural network layer \citep{tai2015improved}.

@@

The TreeLSTM composition function generalizes the LSTM neural network layer to tree- rather than sequence-based inputs, and it shares with the LSTM the idea of representing intermediate states as a pair of a fast-changing state representation $\vec{h}$ and a slower-changing memory representation $\vec{c}$. Our version is formulated as:
\begin{gather}
\colvec{5}
    {\vec{i}}
    {\vec{f}_l}
    {\vec{f}_r}
    {\vec{o}}
    {\vec{g}}
= \colvec{5}
    {\sigma}
    {\sigma}
    {\sigma}
    {\sigma}
    {\text{tanh}}
\left(
W_{\text{comp}}
\colvec{3}
    {\vec{h}_s^1}
    {\vec{h}_s^2}
    {\vec{e}}
+ \vec{b}_{\text{comp}}
\right) \label{eqn:lstm1}
\\
\vec{c} = \vec{f}_l \odot \vec{c}_s^{\,2} + \vec{f}_r \odot \vec{c}_s^{\,1} + \vec{i} \odot \vec{g}  
\\
\vec{h} = \vec{o} \odot \vec{c}
\end{gather}
$\sigma$ is the sigmoid activation function. The results of this function are the pair $\langle\vec{h}, \vec{c}\rangle$, which are placed back on the stack. The two input tree nodes popped off the stack are represented as the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$. In addition, $\vec{e}$ is an optional input argument which is either the empty vector or a vector from an external source like the tracking LSTM (see Section~\ref{sec:tracking}). $\odot$ denotes the elementwise product. Each vector-valued variable listed here is of the same dimension $D$ except $\vec{e}$, which is of the independent dimension $D_t$.

@@

The stack and the buffer can be represented as arrays of $N$ elements each (for sentences of up to $N$ words), with the two $D$-dimensional vectors $\vec{h}$ and $\vec{c}$ in each element.

@@

We use word representations based on the standard 300-dimensional vector package provided with GloVe \citep{pennington2014glove}. We do not update these representations during training. Instead, we use a learned linear transformation to transform each input word vector $\vec{x}_{\text{GloVe}}$ into a pair of vectors $\langle \vec{h}, \vec{c}\rangle$ that are stored side-by-side in the buffer and used as inputs to the TreeLSTM composition function:
\begin{equation}
\colvec{2}
    {\vec{h}}
    {\vec{c}}
= W_{\text{wd}} \vec{x}_{\text{GloVe}} + \vec{b}_{\text{wd}}
\end{equation}

This paper evaluates SPINN on a sentence pair classification task. To classify a sentence pair, we first construct a feature vector $\vec{x}_{\text{classifier}}$ based on the SPINN's representation for each of the two sentences---the value at the head of the stacks after the final transition. In particular, the $\vec{h}$ portions of these representations are used. The final feature vector consists of the concatenation of these two vectors, $\vec{h}_{\text{premise}}$ and $\vec{h}_{\text{hypothesis}}$, their difference, and their elementwise product:
\begin{equation}
\vec{x}_{\text{classifier}} = 
\colvec{4}
    {\vec{h}_{\text{premise}}}
    {\vec{h}_{\text{hypothesis}}}
    {\vec{h}_{\text{premise}} - \vec{h}_{\text{hypothesis}}}
    {\vec{h}_{\text{premise}} \odot \vec{h}_{\text{hypothesis}}}
\end{equation}
Following \citet{snli:emnlp2015}, this feature vector is then passed to a series of ReLU neural network layers (or MLP; the number of layers is tuned as a hyperparameter), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.

@@

In addition to the stack, the buffer, and the composition function, our full model includes an additional component: the tracking LSTM. This is a simple low-dimensional sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step. It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, and it is used for two purposes: it supplies feature representations to the transition classifier, which allows the model to stand alone as a parser, and it additionally supplies a secondary input $\vec{e}$ (see Equation~\ref{eqn:lstm1}) to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.

The tracking LSTM's inputs (highlighted in yellow in Figure~\ref{fig:m1-views}) are: the top element of the buffer $\vec{h}_b^1$, which would be moved in a \shift~operation, and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$, which would be composed in a \reduce~operation.

\paragraph{Why a tree-sequence hybrid?} Lexical ambiguity is ubiquitous in natural language---most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence. Simpler sequence-based sentence encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which acts as a cue for disambiguation. In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with, which is often just a single additional word. Feeding a context representation from the tracking LSTM into the composition function is a simple and efficient way to mitigate this disadvantage of tree-structured models.

@@

For SPINN to be able to operate on unparsed inputs, it needs to be able to produce its own transition sequence $\mathbf t$ rather than relying on an external parser to supply this as part of the input. To do this, the model predicts $t_t$ at each step using a simple two-way softmax classifier over tracking LSTM features:
\begin{equation}
\vec{p}_{\text{t}} = \text{softmax}(W_{\text{trans}}\vec{h}_{\text{tracking}} + \vec{b}_{\text{trans}})
\end{equation}
In the full SPINN model, we follow whichever transition (i.e., \shift~or \reduce) has a higher predicted probability in this distribution. This prediction function is trained to mimic the decisions of an external parser. In particular, we use the transition sequences corresponding to Stanford PCFG Parser constituency parses included with SNLI.

At training time, the model follows the transitions corresponding to the gold parse. At test time, a transition is chosen at each step by taking a hard max over the output of this function. We did not find the technique of scheduled sampling \citep{bengio2015scheduled}, or allowing the model to use its own transition decisions in some instances at training time, to be helpful.

\subsection{Implementing SPINN}

\paragraph{The objective function} Our objective function combines a cross-entropy objective $\mathcal{L}_{\text{s}}$ for the SNLI classification task, a cross-entropy objective $\{\mathcal{L}_p^t, \mathcal{L}_h^t\}$ for each parsing decision for each of the sentence and the hypothesis sentence at each step $t$, and an L2 regularization term on the trained parameters. The terms are weighted using the tuned hyperparameters $\alpha$ and $\lambda$:
\begin{equation}
\begin{split}
\mathcal{L}_{\text{m}} = &\mathcal{L}_{\text{s}} + \alpha \sum_{t=0}^{T-1} (\mathcal{L}_{\text{p}}^{t} + \mathcal{L}_{\text{h}}^{t}) + \lambda \|\theta\|^2_2
\end{split}
\end{equation}
\paragraph{Preparing the data} At training time, SPINN requires both a transition sequence $\mathbf t$  and a token sequence $\mathbf x$ as its inputs for each sentence. The token sequence is simply the words in the sentence in order. $\mathbf t$ can be obtained from any constituency parse for the sentence by first converting that parse into an unlabeled binary parse, then linearizing it (with the usual in-order traversal), then taking each word token as a \shift~transition and each `)' as a \reduce~transition, as here:

@@

@@

We use random search to tune the hyperparameters of the model, setting the ranges for search for each hyperparameter heuristically (and validating the reasonableness of the ranges on the development set), and than launching eight copies of each experiment each with newly sampled hyperparameters from those ranges. Appendix~\ref{sec:hyperparams} shows the precise values used on the strongest models.

@@

\subsection{Efficient stack representation}

A na\"ive implementation of SPINN would require representing a stack of size $N$ for each timestep of each input sentence at training time to support backpropagation. This implies a per-example space requirement of $N \times T \times D$, which is prohibitively large for significant batch sizes or sentence lengths $N$. (Recall that $T = 2N - 1$ is the number of transitions in the shift-reduce parsing sequence, and $D$ is the dimension of the vectors on the stack.) Such a naive implementation would also require duplicating a largely unchanged stack at each timestep, since \shift~and \reduce~operations only affect the top of the stack.

In its bare form, before the addition of the tracking LSTM, SPINN (in particular the SPINN-PI-NT variant, for \textit{parsed input, no tracking}) is precisely equivalent to a conventional tree-structured neural network model in the function that it computes, and it therefore also has the same learning dynamics. In both, the representation of each sentence consists of the representation of the words, combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function). The SPINN, however, is dramatically faster, and supports both integrated parsing and a novel approach to context information.

\begin{table}[t]
\centering
\begin{tabular}{c|cl}
  \toprule
  $t$ & $S$[$t$] & $Q_t$ \\
  \midrule
  0 & $a$ & 0 \\
  1 & $b$ & 0 1 \\
  2 & $c$ & 0 1 2 \\
  3 & $(c~b)$ & 0 3 \\
  4 & $((c~b)~a)$ & 4 \\
  \bottomrule
\end{tabular}
\caption{An example of the thin-stack algorithm computing a \shift-\shift-\shift-\reduce-\reduce~sequence on the input sentence $(a, b, c)$. $S$ is shown in the second column and may be thought of as a list of the tops of the stack at all timesteps $t$. The last two elements of $Q$ specify the rows $t$ which should be involved in a \reduce~operation at the next timestep.}
\label{tbl:thin-stack}
\end{table}

\section{Experiments}

\begin{table*}[t]
  \centering\small
  \begin{tabular}{lrrrr} 
    \toprule
Model                   & Params.    & Trans. acc. (\%)  &   Train acc. (\%)  &   Test acc. (\%) \\
\midrule
\multicolumn{5}{c}{\textbf{Previous non-NN results}}\\
Lexicalized classifier \citep{snli:emnlp2015}
                        & --                & --                    &   99.7   &   78.2      \\
\midrule
\multicolumn{5}{c}{\textbf{Previous sentence encoder-based NN results}}\\
100D LSTM encoders \citep{snli:emnlp2015}
                        & 221k               & --               &   84.8   &   77.6      \\
1024D pretrained GRU encoders \citep{DBLP:journals/corr/VendrovKFU15}
                        & 15m                & --              &   98.8   &   81.4       \\
300D Tree-based CNN encoders \citep{mou2015recognizing}
                        & 3.5m                & --             &   83.4   &   82.1       \\
\midrule
\multicolumn{5}{c}{\textbf{Our results}}\\
300D LSTM RNN encoders         & 3.0m                  & --                &   83.9      &   80.6       \\
300D SPINN-PI-NT (\ii{parsed input, no tracking}) encoders
                        & 3.4m                  & --                &   84.4      &   80.9       \\
300D SPINN-PI (\ii{parsed input}) encoders
                        & 3.7m                  & --                &   89.2      &   \textbf{83.2}       \\
300D SPINN (unparsed input) encoders
                        & 2.7m                  & \result{}           &   87.2    &   82.6      \\          
% \result{300D SPINN-PI, sequence-based attn.   }     
%                         & ?                  & --                &   ?      &   86.3dv         \\
% \result{300D SPINN-PI, tree-based attn. }           
%                         & ?                  & --                &   ?      &   \textbf{?}\\
% \midrule
% \multicolumn{5}{c}{Other previous NN results}\\
% \midrule
% 100D LSTM w/ word-by-word attention \citep{rocktaschel2015reasoning}
%                         & 252k               & --              &   85.3   &   83.5       \\
% 300D mLSTM word-by-word attention model \citep{DBLP:journals/corr/WangJ15b}
%                         & 1.9m               & --             &   92.0   &   86.1      \\
% 300D LSTMN with deep attention fusion \citep{cheng2016long}
%                         & 1.4m               & --                &   92.3   &   \textbf{89.1}      \\
    \bottomrule
  \end{tabular}
\caption{\protect\label{tab:results}Results on SNLI 3-way inference classification. Params. is the approximate number of trained parameters (excluding word embeddings for models where they are trained). Trans. acc. is the model's accuracy in predicting parsing transitions. Train and test are SNLI classification accuracy.} 
\end{table*}

\subsection{Natural language inference and SNLI}

>>>>

A wide range of current models in NLP are built around a neural network component that produces vector representations of sentence meaning \citep{tai2015improved,sutskever2014sequence}. This component, the sentence encoder, is generally formulated as a learned parametric function from a sequence of word vectors to a sentence vector, and this function can take a range of different forms. Common sentence encoders include sequence-based recurrent neural network models (RNNs, see Figure~\ref{fig:batching:good}) with Long Short-Term Memory \citep[LSTM,][]{hochreiter1997long}, which accumulate information over the sentence sequentially, convolutional neural networks over words or characters \citep{kalchbrenner2014convolutional,DBLP:journals/corr/ZhangZL15}, which accumulate information using filters over short local sequences of words, and tree-structured recursive neural networks \citep[TreeRNNs,][see Figure~\ref{fig:batching:bad}]{goller1996learning,socher2011parsing}, which propagate information up a binary parse tree.

Of these, the TreeRNN is the principled choice, since meaning in natural language sentences is constructed incrementally according to a tree structure \citep[][i.a.]{Szabolcsi:2009}. However, while these models have shown promise \citep{tai2015improved,li2015tree,bowman2015trees}, they have largely been overlooked in favor of sequence-based RNNs because of their incompatibility with batched computation and their reliance on external parsers.  Batched computation---performing synchronized computation across many examples at once---yields order-of-magnitude improvements in model run time, and is a crucial enabling technology in allowing neural networks to be trained efficiently on large datasets. Because TreeRNNs use a different model structure for each sentence, batched computation is impossible in standard implementations. In addition, standard TreeRNN models can only operate on sentences that have already been processed by a syntactic parser, which slows and complicates the use of these models at test time for most applications.

This paper introduces a new model to address both these issues: the Stack-augmented Parser-Interpreter Neural Network, or SPINN, shown in Figure~\ref{fig:m1-views}. SPINN executes the computations of a tree-structured model in a linearized sequence, and can incorporate a neural network parser that produces can produce the required parse structure on-the-fly. This design improves upon the TreeRNN architecture in three ways: At test time, it can simultaneously parse and interpret unparsed sentences without incurring a substantial additional computational cost, removing the dependence on an external parser. In addition, it supports batched computation for both parsed and unparsed sentences, supporting dramatic speedups over standard TreeRNNs. Finally, it supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over pure sequence- or tree-based models.

@@

\citet{socher2011parsing,socher2011semi} present versions of the TreeRNN model which are capable of operating over unparsed inputs. However, these methods require an expensive search process at test time. Our model presents a fast alternative approach.

@@

SPINN is inspired by the shift-reduce parsing formalism \citep{aho1972theory}, which builds a tree structure over a sequence (e.g., a natural language sentence) by a single left-to-right scan over its tokens. Shift-reduce parsing is frequently used in natural language parsing \citep[e.g.,][]{nivre2003efficient}.

A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_1, \dots, x_N)$ and uses transitions $\mathbf t = (t_0, \dots, t_{T-1})$, where each $t_t \in \{\shift, \reduce\}$ specifies a \shift~or \reduce~transition (described below). In general a parser may also generate these predictions on-the-fly as it reads the tokens. It reads left-to-right over the input transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure. For any binary-branching tree structure over $N$ words, this requires $T = 2N - 1$ transitions.

The parser uses two auxiliary data structures, a stack $S$ of partially completed subtrees and a buffer $B$ of tokens yet to be parsed. The parser is initialized with the stack empty and with the buffer containing the tokens $\mathbf x$ of the sentence in order. Let $\langle S, B \rangle = \langle \emptyset, \mathbf x \rangle$ denote this starting state. We next proceed through the transition sequence, where each transition $t_t$ selects one of the two following operations. Below the $\mid$ symbol denotes the \textit{cons} (concatenation) operator. We arbitrarily choose to always \textit{cons} on the left in the notation below.
\begin{description}
  \item[\shift:] $\langle S, x \mid B \rangle \to \langle x \mid S, B \rangle$. This operation pops an element from the buffer and pushes it onto the top of the stack.
  \item[\reduce:] $\langle x \mid y \mid S, B \rangle \to \langle (x, y) \mid S, B \rangle$. This operation pops the top two elements from the stack, merges them into a binary tree with children $(x, y)$, and pushes the result back onto the stack.
\end{description}

@@

The core SPINN model implements a sequential shift-reduce parser which operates over sentences. It is designed to produce a vector representation of the sentence as output, as opposed to a tree as in standard shift-reduce parsing. This relies on two key modifications to the shift-reduce algorithm. First, the parser's intermediate representations, including those of partial tree structures on the stack, are fixed-length vectors. In addition, the \reduce~operation combines two such vectors from the stack into another vector using a neural network function.

@@

The TreeLSTM composition function \citep{tai2015improved} generalizes the LSTM neural network layer to tree- rather than sequence-based inputs, and it shares with the LSTM the idea of representing intermediate states as a pair of a fast-changing state representation $\vec{h}$ and a slower-changing memory representation $\vec{c}$. Our version is formulated as:
\begin{gather}
\colvec{5}
    {\vec{i}}
    {\vec{f}_l}
    {\vec{f}_r}
    {\vec{o}}
    {\vec{g}}
= \colvec{5}
    {\sigma}
    {\sigma}
    {\sigma}
    {\sigma}
    {\text{tanh}}
\left(
W_{\text{comp}}
\colvec{3}
    {\vec{h}_s^1}
    {\vec{h}_s^2}
    {\vec{e}}
+ \vec{b}_{\text{comp}}
\right) \label{eqn:lstm1}
\\
\vec{c} = \vec{f}_l \odot \vec{c}_s^{\,2} + \vec{f}_r \odot \vec{c}_s^{\,1} + \vec{i} \odot \vec{g}  
\\
\vec{h} = \vec{o} \odot \vec{c}
\end{gather}
where $\sigma$ is the sigmoid activation function and $\odot$ is the elementwise product. The two input tree nodes popped off the stack are represented as the pairs $\langle\vec{h}^1_s, \vec{c}^{\,1}_s\rangle$ and $\langle\vec{h}^2_s, \vec{c}^{\,2}_s\rangle$. In addition, $\vec{e}$ is an optional input argument which is either the empty vector or a vector from an external source like the tracking LSTM (see Section~\ref{sec:tracking}). The result of this function is the pair $\langle\vec{h}, \vec{c}\rangle$, which is placed back on the stack. Each vector-valued variable listed here is of the same dimension $D$ except $\vec{e}$, which is of the independent dimension $D_{\text{tracking}}$.

@@

The stack and the buffer are arrays of $N$ elements each (for sentences of up to $N$ words), with the two $D$-dimensional vectors $\vec{h}$ and $\vec{c}$ in each element.

@@

We use word representations based on the standard 300D vector package provided with GloVe \citep{pennington2014glove}. We do not update these representations during training. Instead, we use a learned linear transformation to transform each input word vector $\vec{x}_{\text{GloVe}}$ into a pair of vectors $\langle \vec{h}, \vec{c}\rangle$ that are stored side-by-side in the buffer and used as inputs to the TreeLSTM composition function:
\begin{equation}
\colvec{2}
    {\vec{h}}
    {\vec{c}}
= W_{\text{wd}} \vec{x}_{\text{GloVe}} + \vec{b}_{\text{wd}}
\end{equation}

@@

In addition to the stack, the buffer, and the composition function, our full model includes an additional component: the tracking LSTM. This is a simple low-dimensional sequence-based LSTM RNN that operates in tandem with the model, taking inputs from the buffer and stack at each step. It is meant to maintain a low-resolution summary of the portion of the sentence that has been processed so far, which is used for two purposes: it supplies feature representations to the transition classifier, which allows the model to stand alone as a parser, and it additionally supplies a secondary input $\vec{e}$ (see Equation~\ref{eqn:lstm1}) to the composition function, allowing context information to leak into the construction of sentence meaning, and forming what is effectively a tree-sequence hybrid model.

The tracking LSTM's inputs (highlighted in yellow in Figure~\ref{fig:m1-views}) are the top element of the buffer $\vec{h}_b^1$ (which would be moved in a \shift~operation) and the top two elements of the stack $\vec{h}_s^1$ and $\vec{h}_s^2$ (which would be composed in a \reduce~operation).

\paragraph{Why a tree-sequence hybrid?}

Lexical ambiguity is ubiquitous in natural language---most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence. Simpler sequence-based sentence encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which acts as a cue for disambiguation. In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with, which is often just a single additional word. Feeding a context representation from the tracking LSTM into the composition function is a simple and efficient way to mitigate this disadvantage of tree-structured models.

@@

For SPINN to be able to operate on unparsed inputs, it needs to be able to produce its own transition sequence $\mathbf t$ rather than relying on an external parser to supply is as part of the input. To do this, the model predicts $t_t$ at each step using a simple two-way softmax classifier over tracking LSTM features:
\begin{equation}
\vec{p}_{\text{t}} = \text{softmax}(W_{\text{trans}}\vec{h}_{\text{tracking}} + \vec{b}_{\text{trans}})
\end{equation}
At test time, the model uses whichever transition (i.e., \shift~or \reduce) is assigned a higher probability. The prediction function is trained to mimic the decisions of an external parser, and the decisions of this external parser are used as the inputs to the model during training. For SNLI, we use the binary Stanford PCFG Parser parses that are included with the corpus for this purpose. We did not find the technique of scheduled sampling \citep{bengio2015scheduled}, or allowing the model to use its own transition decisions in some instances at training time, to be helpful.

\subsection{Implementation issues}

\paragraph{Representing the stack efficiently}

A na\"ive implementation of SPINN would require representing a stack of size $N$ for each timestep of each input sentence at training time to support backpropagation. This implies a per-example space requirement of $N \times T \times D$, which is prohibitively large for significant batch sizes or sentence lengths $N$. Such a naive implementation would also require copying a largely unchanged stack at each timestep, since \shift~and \reduce~operations only affect the top of the stack.

\paragraph{Preparing the data} At training time, SPINN requires both a transition sequence $\mathbf t$  and a token sequence $\mathbf x$ as its inputs for each sentence. The token sequence is simply the words in the sentence in order. $\mathbf t$ can be obtained from any constituency parse for the sentence by first converting that parse into an unlabeled binary parse, then linearizing it (with the usual in-order traversal), then taking each word token as a \shift~transition and each `)' as a \reduce~transition, as here:

@@

Without the addition of the tracking LSTM, SPINN (in particular the SPINN-PI-NT variant, for \textit{parsed input, no tracking}) is precisely equivalent to a conventional tree-structured neural network model in the function that it computes, and it therefore also has the same learning dynamics. In both, the representation of each sentence consists of the representation of the words, combined recursively using a TreeRNN composition function (in our case, the TreeLSTM function). The SPINN, however, is dramatically faster, and supports both integrated parsing and a novel approach to context through the tracking LSTM.

\begin{table}[t]
\centering
\begin{tabular}{c|cl}
  \toprule
  $t$ & $S$[$t$] & $Q_t$ \\
  \midrule
  0 & $a$ & \underline{\hphantom{0} 0} \\
  1 & $b$ & \hphantom{0} \underline{0 1} \\
  2 & $c$ & \hphantom{0} 0 \underline{1 2} \\
  3 & $(c~b)$ & \hphantom{0} \underline{0 3} \\
  4 & $((c~b)~a)$ & \underline{\hphantom{0} 4} \\
  \bottomrule
\end{tabular}
\caption{An example of the thin-stack algorithm computing a \shift-\shift-\shift-\reduce-\reduce~sequence on the input sentence $(a, b, c)$. $S$ is shown in the second column and may be thought of as a list of the tops of the stack at all timesteps $t$. The last two elements of $Q$ (underlined) specify which rows $t$ would be involved in a \reduce~operation at the next timestep.}
\label{tbl:thin-stack}
\end{table}

In this section, we compare the test time speed of our SPINN-PI-NT with an equivalent TreeRNN implemented in the conventional fashion. While the full models evaluated below are implemented and trained using Theano \citep{bergstra+al:2010-scipy,Bastien-Theano-2012}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of both models here. To do this, we reimplement the feedforward of SPINN-PI-NT in C++/CUDA, and we compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as our model.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}. We plan to release our modified code at publication time, alongside the optimized C++/CUDA SPINN and the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the feedforward/inference pass, as inference speed is the relevant performance metric for most practical applications.}

Each model is restricted to run on sentences of 30 tokens or fewer. We fix the model dimension and the word embedding dimension at 300. We run the CPU performance test on a 2.20-GHz Intel Xeon processor with hyperthreading enabled. Our GPU model is implemented in C++ and CUDA. We test it on a NVIDIA Titan X GPU.

Figure~\ref{fig:speed} compares the sentence encoding speed of the two models using data from the Stanford Sentiment Treebank \citep{socher2013recursive}. Figure~\ref{fig:speed} shows a substantial difference in runtime that increases with batch size, as is typical of GPU computation. At the larger batch sizes that are most effective for test-time applications (e.g., 512+ examples), we observe a 10--20x speedup relative to the standard CPU implementation.

\section{NLI Experiments}

We evaluate SPINN on the task of natural language inference \citep[NLI, a.k.a. recognizing textual entailment, or RTE;][]{dagan2006pascal}. NLI is a sentence pair classification task, in which a model reads two sentences (a premise and a hypothesis), and outputs a judgment of {\it entailment}, {\it contradiction}, or {\it neutral}, reflecting the relationship between the meanings of the two sentences, as in this example from the SNLI corpus, which we use:

\subsection{Applying SPINN to SNLI}

To classify an SNLI sentence pair, we run two copies of SPINN with shared parameters: one on the premise sentence and another the hypothesis sentence. We then use their outputs (the $\vec{h}$ states at the top of each stack at time $t=T$) to construct a feature vector $\vec{x}_{\text{classifier}}$ for the pair. This feature vector consists of the concatenation of these two sentence vectors, their difference, and their elementwise product:
\begin{equation}
\vec{x}_{\text{classifier}} = 
\colvec{4}
    {\vec{h}_{\text{premise}}}
    {\vec{h}_{\text{hypothesis}}}
    {\vec{h}_{\text{premise}} - \vec{h}_{\text{hypothesis}}}
    {\vec{h}_{\text{premise}} \odot \vec{h}_{\text{hypothesis}}}
\end{equation}
Following \citet{snli:emnlp2015}, this feature vector is then passed to a series of ReLU neural network layers (i.e., an MLP; the number of layers is tuned as a hyperparameter), then passed into a linear transformation, and then finally passed to a softmax layer, which yields a distribution among the three labels.

\begin{table*}[t]
  \centering\small
  \begin{tabular}{lrrrr} 
    \toprule
Model                   & Params.    & Trans. acc. (\%)  &   Train acc. (\%)  &   Test acc. (\%) \\
\midrule
\multicolumn{5}{c}{\textbf{Previous non-NN results}}\\
Lexicalized classifier \citep{snli:emnlp2015}
                        & --                & --                    &   99.7   &   78.2      \\
\midrule
\multicolumn{5}{c}{\textbf{Previous sentence encoder-based NN results}}\\
100D LSTM encoders \citep{snli:emnlp2015}
                        & 221k               & --               &   84.8   &   77.6      \\
1024D pretrained GRU encoders \citep{DBLP:journals/corr/VendrovKFU15}
                        & 15m                & --              &   98.8   &   81.4       \\
300D Tree-based CNN encoders \citep{mou2015recognizing}
                        & 3.5m                & --             &   83.4   &   82.1       \\
\midrule
\multicolumn{5}{c}{\textbf{Our results}}\\
300D LSTM RNN encoders         & 3.0m                  & --                &   83.9      &   80.6       \\
300D SPINN-PI-NT (\ii{parsed input, no tracking}) encoders
                        & 3.4m                  & --                &   84.4      &   80.9       \\
300D SPINN-PI (\ii{parsed input}) encoders
                        & 3.7m                  & --                &   89.2      &   \textbf{83.2}       \\
300D SPINN (unparsed input) encoders
                        & 2.7m                  & \result{}           &   87.2    &   82.6      \\          
% \result{300D SPINN-PI, sequence-based attn.   }     
%                         & ?                  & --                &   ?      &   86.3dv         \\
% \result{300D SPINN-PI, tree-based attn. }           
%                         & ?                  & --                &   ?      &   \textbf{?}\\
% \midrule
% \multicolumn{5}{c}{Other previous NN results}\\
% \midrule
% 100D LSTM w/ word-by-word attention \citep{rocktaschel2015reasoning}
%                         & 252k               & --              &   85.3   &   83.5       \\
% 300D mLSTM word-by-word attention model \citep{DBLP:journals/corr/WangJ15b}
%                         & 1.9m               & --             &   92.0   &   86.1      \\
% 300D LSTMN with deep attention fusion \citep{cheng2016long}
%                         & 1.4m               & --                &   92.3   &   \textbf{89.1}      \\
    \bottomrule
  \end{tabular}
\caption{\protect\label{tab:results}Results on SNLI 3-way inference classification. Params. is the approximate number of trained parameters (excluding word embeddings for models where they are trained). Trans. acc. is the model's accuracy in predicting parsing transitions at test time. Train and test are SNLI classification accuracy.} 
\end{table*}

\paragraph{The objective function} Our objective combines a cross-entropy objective $\mathcal{L}_{\text{s}}$ for the SNLI classification task, a cross-entropy objective $\{\mathcal{L}_p^t, \mathcal{L}_h^t\}$ for each parsing decision for each of the sentences at each step $t$, and an L2 regularization term on the trained parameters. The terms are weighted using the tuned hyperparameters $\alpha$ and $\lambda$:
\begin{equation}
\begin{split}
\mathcal{L}_{\text{m}} = &\mathcal{L}_{\text{s}} + \alpha \sum_{t=0}^{T-1} (\mathcal{L}_{\text{p}}^{t} + \mathcal{L}_{\text{h}}^{t}) + \lambda \|\theta\|^2_2
\end{split}
\end{equation}

@@

@@

Appendix~\ref{sec:hyperparams} discusses hyperparameter tuning.
